A Plasma Torus Around a Young Low-Mass Star

### 2025-03-11 paper/ms.tex
update

\begin{table*}
\scriptsize
\setlength{\tabcolsep}{2pt}
\centering
\caption{Selected System Parameters for TIC~141146667}
\label{tab:params}
\begin{tabular}{llcc}
\hline \hline
Parameter & Description & Value & Source\\
\hline 
%
$T_{\rm eff}$\dotfill                   & Effective Temperature (K) \hspace{9pt}\dotfill                 & 2972 $\pm$ 40    & 1 \\
%
%$\log{g_{\star}}$\dotfill              & Surface Gravity (cgs)\hspace{9pt}\dotfill                      & YYY              & X \\
%
$R_\star$\dotfill                       & Stellar radius ($R_\odot$)\dotfill                             & 0.42$\pm$0.02    & 1 \\
%
$M_\star$\dotfill                       & Stellar mass ($M_\odot$)\dotfill                               & XXX  & 6 \\
%
$\gamma$\dotfill                        & Systemic radial velocity (\kms)\dotfill                        & 0.61 $\pm$ 1.47  & 1  \\
%
Age                                     & Adopted stellar age (Myr)\dotfill                              & YYY  & 8 \\
%
Spec. Type\dotfill                      & Spectral Type\dotfill                                          & MXV              & X \\
%
$P_{\rm rot}$\dotfill                   & Rotation period (hr)\dotfill                                   & $3.930\pm 0.001$ & X \\
%
$v_{\rm eq}$\dotfill		                & Equatorial velocity \dotfill                                   &  130$\pm$4       & - \\
                                        & \hspace{3pt} ($2\pi R_\star/P_{\rm rot}$) (\kms)	             &                      \\
%
$v_{\rm eq}\sin{i_\star}$\dotfill		    & Projected rotational\dotfill                                   &  128$\pm$3       & X \\
                                        & \hspace{3pt} velocity (\kms)	                                 &                      \\
%
$i_\star$\dotfill                       & Stellar inclination (deg)\dotfill                              & 	XXX             & X \\
%
$d$\dotfill                             & Distance (pc)\dotfill                                          & $57.6 \pm X.X$   & X \\
%
$R_{\rm c}$\dotfill		                  & Keplerian corotation\dotfill                                   &  XXX             & X \\
                                        & \hspace{3pt} radius ($R_\star$)	                               &                      \\
%
$a_1$\dotfill                           & Clump 1 orbital radius ($R_\star$)\hspace{9pt}\dotfill         &  2.1-2.7         & X \\
$a_2$\dotfill                           & Clump 2 orbital radius ($R_\star$)\hspace{9pt}\dotfill         &  3.9             & X \\
%
$\langle$EW$_{\rm H\alpha}$$\rangle$    & Time-averaged H$\alpha$ line core                              &  X.X             & Y \\ 
                                        & \hspace{3pt} equivalent width (\AA)	                           &                      \\
Range(EW$_{\rm H\alpha}$)$_1$           & H$\alpha$ equiv. width range                                   &  X.X             & Y \\ 
                                        & \hspace{3pt} from Clump 1 (\AA)	                               &                      \\
Range(EW$_{\rm H\alpha}$)$_2$           & H$\alpha$ equiv. width range                                   &  X.X             & Y \\ 
                                        & \hspace{3pt} from Clump 2 (\AA)	                               &                      \\
\hline
\end{tabular}
\begin{flushleft}
\footnotesize{ \textsc{NOTE}---
%$^\dagger$ The GAIADR2 and GAIAEDR3 identifiers for Kepler 1627A are identical.  The secondary
is not resolved in the Gaia point source catalog.
$^*$ Given only $v\sin i$ and $2\pi R_\star/P_{\rm rot}$, $\cos i=0.11^{+0.11}_{-0.08}$.
Provenances are:
1: SED fit\cite{Bouma2024},
%$^6$Cluster isochrone (MIST adopted; PARSEC compared for quoted
%  uncertainty),
2: TESS light curve,
%  3: HIRES spectra and specmatch(?),
%$^8$Pre-main-sequence CAMD interpolation (Section~\ref{sec:camd}),
}
\end{flushleft}
\vspace{-0.5cm}
\end{table*}

@@

\subsection{Stellar Parameters}\phantom{+}

@@

For TIC~141146667, we measured the velocities from each of our HIRES
spectra using regions near the K~I (7700\,\AA) resonance line and
three TiO bandheads (5160\,\AA, 5450\,\AA, and 5600\,\AA).  These
regions were selected because they provided the best matches between
the synthetic and observed spectra.  The resulting redshift
measurements over each order were averaged to produce the final
measurement.  The star's 130$\pm$4\,\kms\ equatorial velocity and
near-edge on viewing orientation yielded a line broadening $\Delta
\lambda$$\approx$3\,\AA.  We used the scatter of resulting velocity
measurements between orders to assign the RV uncertainty at each
epoch.  The uncertainty-weighted mean systemic velocity over all
epochs on UT 2024-02-17 was $\gamma$=0.6$\pm$1.5\,\kms.  The relative
radial velocities about this mean are given in Table~\ref{tab:rv}.
% from calc_max_amplitude_sinusoid_rv_timeseries.py
% from K_to_msini.py, assuming Mstar = 0.18

@@

% TEMPLATE: IRAS041
%
\begin{table}
    \centering
    \begin{tabular}{lcr}
    \hline 
    \hline
    Parameter & Host & Source \\
    \hline 
    \multicolumn{3}{c}{Identifiers} \\
    \hline
    TIC & 141146667 & TESS \\
    Gaia & 860453786736413568 & Gaia\ DR3 \\
    %2MASS & J04154278+2909597 & J04154269+2909558 & 2MASS \\
    %ALLWISE & J041542.77+290959.5 & ... & ALLWISE\\
    \hline
    \multicolumn{3}{c}{Astrometry} \\ 
    \hline
    $\alpha$ & todo & Gaia\ DR3 \\
    $\delta$ & todo & Gaia\ DR3 \\
    $\mu_{\alpha}$ (mas yr$^{-1}$ ) & -73.933 $\pm$ 0.022 & Gaia\ DR3 \\
    $\mu_{\delta}$ (mas yr$^{-1}$ ) &  32.262 $\pm$ 0.024 & Gaia\ DR3 \\
    $\pi$ (mas)                     &  17.324 $\pm$ 0.025 & Gaia\ DR3 \\
    RUWE                            &  1.23               & Gaia\ DR3 \\
    \hline
    \multicolumn{3}{c}{Photometry} \\
    \hline
    $TESS$ (mag) & todo & TESS\ \\
    $G$ (mag)                       & 14.701 $\pm$ 0.002 & Gaia\ DR3 \\
    $G_{\rm BP}$ (mag)              & 16.664 $\pm$ 0.008 & Gaia\ DR3 \\
    $G_{\rm RP}$ (mag)              & 13.398 $\pm$ 0.006 & Gaia\ DR3 \\
    $G_{\rm BP}$-$G_{\rm RP}$ (mag) &  3.276 $\pm$ 0.010 & Gaia\ DR3 \\
    $J$ (mag)                       & 11.401 $\pm$ 0.022 & 2MASS     \\
    $H$ (mag)                       & 10.793 $\pm$ 0.021 & 2MASS     \\
    $K_s$ (mag)                     & 10.473 $\pm$ 0.016 & 2MASS     \\
    $W1$ (mag)                      & 10.276 $\pm$ 0.023 & ALLWISE   \\ % Cutri+2013 II/328/allwise
    $W2$ (mag)                      & 10.070 $\pm$ 0.020 & ALLWISE   \\
    $W3$ (mag)                      &  9.838 $\pm$ 0.045 & ALLWISE   \\
    %$W4$ (mag)                      &  8.392 $\pm$ NOUNC & ALLWISE \\
	  %
    %TODO: eROSITA?  ROSAT is an emitter.  GALEX detection.
    % Ofek+2011 1.4Ghz emitter.  Helfand+2015 ditto.  Bruzewski+2021.
    \hline
    \multicolumn{3}{c}{Kinematics and Position} \\
    \hline
    $RV_{Bary}$ (km s$^{-1}$ ) & $13.35 \pm 3.39$ & HIRES \\
    $U$ (\kms) & & \\
    $V$ (\kms) & & \\
    $W$ (\kms) & & \\
    $X$ (pc) & -28.4 & \\
    $Y$ (pc) &  19.8 & \\
    $Z$ (pc) &  67.0 & \\
    \hline
    \multicolumn{3}{c}{Physical Properties} \\
    \hline
    $P_{rot}$ (hours) & $3.930 \pm 0.XXX$ & This work \\ 
    $v \sin i_\star$(km s$^{-1}$) & todo & This work\\
    $i_\star$($^\circ$) & todo & This work \\
    $T_{eff}$ (K) & 2972 $\pm$ 40 & \cite{Bouma2024}\\
    %$\log{g_{\star}}$\dotfill              & Surface Gravity (cgs)\hspace{9pt}\dotfill                      & YYY              & X \\
    $R_\star$ ($R_{\odot}$) & 0.42 $\pm$ 0.02 & \cite{Bouma2024} \\
    $A_V$ (mag) & 0 & \cite{Green2019} \\
    $F_{bol}$ (erg cm$^{-2}$ s$^{-1}$ ) & todo & This work\\
    $L_\star$ ($L_{\odot}$)  & todo & This work\\
    $M_\star$ ($M_{\odot}$)  & todo & This work\\
    $t_{\rm iso}$ (Myr) & $16^{+19}_{-6}$ &  This work \\
    \hline
    \end{tabular}
    \caption{Properties of \starname.}
    \label{tab:stellarParameters}
\end{table}

>>>>

\begin{table*}
\scriptsize
\setlength{\tabcolsep}{2pt}
\centering
\caption{Selected System Parameters for TIC~141146667}
\label{tab:params}
\begin{tabular}{llcc}
\hline \hline
Parameter & Description & Value & Source\\
\hline 
%
$T_{\rm eff}$\dotfill                   & Effective Temperature (K) \hspace{9pt}\dotfill                 & 2972 $\pm$ 40    & 1 \\
%
%$\log{g_{\star}}$\dotfill              & Surface Gravity (cgs)\hspace{9pt}\dotfill                      & YYY              & X \\
%
$R_\star$\dotfill                       & Stellar radius ($R_\odot$)\dotfill                             & 0.42$\pm$0.02    & 1 \\
%
Age                                     & Adopted stellar age (Myr)\dotfill                              & $16^{+19}_{-6}$  & 2 \\
%
$M_\star$\dotfill                       & Stellar mass ($M_\odot$)\dotfill                               & XXX  &           & 3 \\
%
$\gamma$\dotfill                        & Systemic radial velocity (\kms)\dotfill                        & 0.61 $\pm$ 1.47  & 4 \\
%
Spec. Type\dotfill                      & Spectral Type\dotfill                                          & MXV              & 4 \\ % TODO TODO
%
$P_{\rm rot}$\dotfill                   & Rotation period (hr)\dotfill                                   & $3.930\pm 0.001$ & 5 \\
%
$v_{\rm eq}$\dotfill		                & Equatorial velocity \dotfill                                   &  130$\pm$4       & 6 \\
                                        & \hspace{3pt} ($2\pi R_\star/P_{\rm rot}$) (\kms)	             &                      \\
%
$v_{\rm eq}\sin{i_\star}$\dotfill		    & Projected rotational\dotfill                                   &  138$\pm$8       & 4 \\
                                        & \hspace{3pt} velocity (\kms)	                                 &                      \\
%
$i_\star$\dotfill                       & Stellar inclination\dotfill                                    & 	$>$63           & 4 \\
                                        & \hspace{3pt}  2$\sigma$ lower limit (deg)	                     &                      \\
%
$d$\dotfill                             & Distance (pc)\dotfill                                          & $57.6 \pm X.X$   & X \\
%
$R_{\rm c}$\dotfill		                  & Keplerian corotation\dotfill                                   &  XXX             & X \\
                                        & \hspace{3pt} radius ($R_\star$)	                               &                      \\
%
$a_1$\dotfill                           & Clump 1 orbital radius ($R_\star$)\hspace{9pt}\dotfill         &  2.1-2.7         & X \\
$a_2$\dotfill                           & Clump 2 orbital radius ($R_\star$)\hspace{9pt}\dotfill         &  3.9             & X \\
%
$\langle$EW$_{\rm H\alpha}$$\rangle$    & Time-averaged H$\alpha$ line core                              &  X.X             & Y \\ 
                                        & \hspace{3pt} equivalent width (\AA)	                           &                      \\
Range(EW$_{\rm H\alpha}$)$_1$           & H$\alpha$ equiv. width range                                   &  X.X             & Y \\ 
                                        & \hspace{3pt} from Clump 1 (\AA)	                               &                      \\
Range(EW$_{\rm H\alpha}$)$_2$           & H$\alpha$ equiv. width range                                   &  X.X             & Y \\ 
                                        & \hspace{3pt} from Clump 2 (\AA)	                               &                      \\
\hline
\end{tabular}
\begin{flushleft}
\footnotesize{ \textsc{NOTE}---
$^*$ Given only $v\sin i$ and $2\pi R_\star/P_{\rm rot}$, $\cos i=0.11^{+0.11}_{-0.08}$.
Provenances are:
1: SED fit\cite{Bouma2024},
2: Gaia DR3 empirical isochrone fit (see~\ref{subsec:stparams}),
3: PARSEC %(TODO)
%$^6$Cluster isochrone (MIST adopted; PARSEC compared for quoted
%  uncertainty),
4: Keck/HIRES.
5: TESS light curve,
6: Derived quantity.
%  3: HIRES spectra and specmatch(?),
%$^8$Pre-main-sequence CAMD interpolation (Section~\ref{sec:camd}),
}
\end{flushleft}
\vspace{-0.5cm}
\end{table*}

@@

\subsection{Stellar Parameters}\phantom{+}
\label{subsec:stparams}

@@

For TIC~141146667, we measured the radial velocities from each of our
HIRES spectra using regions near the K~I (7700\,\AA) resonance line and
three TiO bandheads (5160\,\AA, 5450\,\AA, and 5600\,\AA).  These
regions were selected because they provided the best matches between the
synthetic and observed spectra.  The resulting redshift measurements
over each order were averaged to produce the final measurement.  We used
the scatter of resulting velocity measurements between orders to assign
the RV uncertainty at each epoch.  The uncertainty-weighted mean
systemic velocity over all epochs on UT 2024-02-17 was
$\gamma$=0.6$\pm$1.5\,\kms.  The relative radial velocities about this
mean are given in Table~\ref{tab:rv}.

{\it Viewing Orientation}---We fitted the rotational broadening of the
K~I (7700\,\AA) resonance line using the broadening kernel suggested by
Ref.~\cite{Gray2008}; taking the mean and standard deviation of the
result value over all epochs yielded $v_{\rm eq} \sin
i$=138$\pm$8\,\kms, consistent with the visual line broadening $\Delta
\lambda$$\approx$3\,\AA.  The star's equatorial velocity $v_{\rm eq}$
based on its apparent size and rotation period is 130$\pm$4\,\kms.
While this suggests that the viewing orientation could be nearly
edge-on, the formal constraint is rather weak, with $i$$>$63$^\circ$ at
2$\sigma$.
% from calc_max_amplitude_sinusoid_rv_timeseries.py
% from K_to_msini.py, assuming Mstar = 0.18

@@

% TEMPLATE: IRAS041
%
\begin{table}
    \centering
    \begin{tabular}{lcr}
    \hline 
    \hline
    Parameter & Host & Source \\
    \hline 
    \multicolumn{3}{c}{Identifiers} \\
    \hline
    TIC & 141146667 & TESS \\
    Gaia & 860453786736413568 & Gaia\ DR3 \\
    %2MASS & J04154278+2909597 & J04154269+2909558 & 2MASS \\
    %ALLWISE & J041542.77+290959.5 & ... & ALLWISE\\
    \hline
    \multicolumn{3}{c}{Astrometry} \\ 
    \hline
    $\alpha$ & todo & Gaia\ DR3 \\
    $\delta$ & todo & Gaia\ DR3 \\
    $\mu_{\alpha}$ (mas yr$^{-1}$ ) & -73.933 $\pm$ 0.022 & Gaia\ DR3 \\
    $\mu_{\delta}$ (mas yr$^{-1}$ ) &  32.262 $\pm$ 0.024 & Gaia\ DR3 \\
    $\pi$ (mas)                     &  17.324 $\pm$ 0.025 & Gaia\ DR3 \\
    RUWE                            &  1.23               & Gaia\ DR3 \\
    \hline
    \multicolumn{3}{c}{Photometry} \\
    \hline
    $TESS$ (mag) & todo & TESS\ \\
    $G$ (mag)                       & 14.701 $\pm$ 0.002 & Gaia\ DR3 \\
    $G_{\rm BP}$ (mag)              & 16.664 $\pm$ 0.008 & Gaia\ DR3 \\
    $G_{\rm RP}$ (mag)              & 13.398 $\pm$ 0.006 & Gaia\ DR3 \\
    $G_{\rm BP}$-$G_{\rm RP}$ (mag) &  3.276 $\pm$ 0.010 & Gaia\ DR3 \\
    $J$ (mag)                       & 11.401 $\pm$ 0.022 & 2MASS     \\
    $H$ (mag)                       & 10.793 $\pm$ 0.021 & 2MASS     \\
    $K_s$ (mag)                     & 10.473 $\pm$ 0.016 & 2MASS     \\
    $W1$ (mag)                      & 10.276 $\pm$ 0.023 & ALLWISE   \\ % Cutri+2013 II/328/allwise
    $W2$ (mag)                      & 10.070 $\pm$ 0.020 & ALLWISE   \\
    $W3$ (mag)                      &  9.838 $\pm$ 0.045 & ALLWISE   \\
    %$W4$ (mag)                      &  8.392 $\pm$ NOUNC & ALLWISE \\
	  %
    %TODO: eROSITA?  ROSAT is an emitter.  GALEX detection.
    % Ofek+2011 1.4Ghz emitter.  Helfand+2015 ditto.  Bruzewski+2021.
    \hline
    \multicolumn{3}{c}{Kinematics and Position} \\
    \hline
    $RV_{Bary}$ (km s$^{-1}$ ) & $13.35 \pm 3.39$ & HIRES \\
    $U$ (\kms) & & \\
    $V$ (\kms) & & \\
    $W$ (\kms) & & \\
    $X$ (pc) & -28.4 & \\
    $Y$ (pc) &  19.8 & \\
    $Z$ (pc) &  67.0 & \\
    \hline
    \multicolumn{3}{c}{Physical Properties} \\
    \hline
    $P_{rot}$ (hours) & $3.930 \pm 0.XXX$ & This work \\ 
    $v \sin i_\star$(km s$^{-1}$) & $138 \pm 8$ & This work\\
    $i_\star$($^\circ$) & $>$63 & This work \\
    $T_{eff}$ (K) & 2972 $\pm$ 40 & \cite{Bouma2024}\\
    %$\log{g_{\star}}$\dotfill              & Surface Gravity (cgs)\hspace{9pt}\dotfill                      & YYY              & X \\
    $R_\star$ ($R_{\odot}$) & 0.42 $\pm$ 0.02 & \cite{Bouma2024} \\
    $A_V$ (mag) & 0 & \cite{Green2019} \\
    $F_{bol}$ (erg cm$^{-2}$ s$^{-1}$ ) & todo & This work\\
    $L_\star$ ($L_{\odot}$)  & todo & This work\\
    $M_\star$ ($M_{\odot}$)  & todo & This work\\
    $t_{\rm iso}$ (Myr) & $16^{+19}_{-6}$ &  This work \\
    \hline
    \end{tabular}
    \caption{Properties of \starname.}
    \label{tab:stellarParameters}
\end{table}
